---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# olsrr

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/olsrr)](https://cran.r-project.org/package=olsrr)
[![Travis-CI Build Status](https://travis-ci.org/rsquaredacademy/olsrr.svg?branch=master)](https://travis-ci.org/rsquaredacademy/olsrr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/rsquaredacademy/olsrr?branch=master&svg=true)](https://ci.appveyor.com/project/rsquaredacademy/olsrr)
[![Coverage Status](https://img.shields.io/codecov/c/github/rsquaredacademy/olsrr/master.svg)](https://codecov.io/github/rsquaredacademy/olsrr?branch=master)

## Overview

The olsrr package provides following tools for teaching and learning OLS regression using R:

- Comprehensive Regression Output

- Variable Selection Procedures
	- All Possible Regression
	- Best Subsets Regression
	- Stepwise Regression
	- Stepwise Forward Regression
	- Stepwise Backward Regression
	- stepAIC Regression
	- stepAIC Forward Regression
	- stepAIC Backward Regression

- Heteroskedasticity Tests
	- Bartlett Test
	- Breusch Pagan Test
	- F Test
	- Score Test

- Collinearity Diagnostics
	- Variance Inflation Factors
	- Tolerance
	- Eigenvalues
	- Condition Indices

- Model Fit Assessment
	- Residual Fit Spread Plot
	- Part & Partial Correlations
	- Observed vs Fitted Values Plot
	- Lack of Fit F Test
	- Diagnostics Panel

- Measures of Influence
	- Cook's D Bar Plot
	- Cook's D Chart
	- DFBETAs Panel
	- DFFITS Plot
	- Studentized Residual Plot
	- Studentized Residual Chart
	- Studentized Residuals vs Leverage Plot
	- Deleted Studentized Residuals vs Fitted Values Plot
	- Hadi Plot
	- Potential Residual Plot

- Residual Diagnostics
	- Residual QQ Plot
	- Residual Histogram
	- Residual Box Plot
	- Residual Normality Test
	- Residual vs Fitted Values Plot

- Variable Contribution Assessment
	- Residual vs Regressor Plot
	- Added Variable Plot
	- Residual Plus Component Plot


## Installation

You can install olsrr from github with:

```{r cran-installation, eval = FALSE}
# install olsrr from CRAN
install.packages("olsrr")

# the development version from github
# install.packages("devtools")
devtools::install_github("rsquaredacademy/olsrr")
```

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
